---
title: "Hands-on Exercise 1"
author: "Andre Ong Jia Kang"
---

# Getting Started

## Install and launching R packages

The code chunk below uses p_load() of pacman package to check if tidyverse packages are installed in the computer. If they are, then they will be launched into R.

```{r}
pacman::p_load(ggrepel, patchwork, 
               ggthemes, hrbrthemes,
               tidyverse, ggiraph, plotly, DT) 
```

## Importing the data

```{r}
exam_data <- read_csv("data/Exam_data.csv")

```

## Plotting a simple bar chart

```{r}
ggplot(data = exam_data, aes(x = RACE)) +
  geom_bar()
```

# 1.ggplot2 methods

## 1.3 Introducing ggplot

### 1.3.1 R Graphics VS ggplot

```{r}
ggplot(data=exam_data, aes(x = MATHS)) +
  geom_histogram(bins=10, boundary = 100, color="black", fill="grey") +
  ggtitle("Distribution of Maths scores")
```

## 1.4 Grammar of Graphics

In the nutshell, Grammar of Graphics defines the rules of structuring mathematical and aesthetic elements into a meaningful graph.

There are two principles in Grammar of Graphics, they are:

-   Graphics = distinct layers of grammatical elements

-   Meaningful plots through aesthetic mapping

### 1.4.1 A Layered Grammaer of Graphics

-   A short description of each building block are as follows:

-   Data: The dataset being plotted.

-   Aesthetics take attributes of the data and use them to influence visual characteristics, such as position, colours, size, shape, or transparency.

-   Geometrics: The visual elements used for our data, such as point, bar or line.

-   Facets split the data into subsets to create multiple variations of the same graph (paneling, multiple plots).

-   Statistics, statiscal transformations that summarise data (e.g. mean, confidence intervals).

-   Coordinate systems define the plane on which data are mapped on the graphic.

-   Themes modify all non-data components of a plot, such as main title, sub-title, y-aixs title, or legend background.

## 1.5 Essential Grammatical Elements in ggplot2: data
```{r}
ggplot(data=exam_data)
```
## 1.6 Essential Grammatical Elements in ggplot2: data
```{r}
ggplot(data=exam_data, 
       aes(x= MATHS))
```
## 1.7 Essential Grammatical Elements in ggplot2: Aesthetic mappings

### 1.7.1 Geometric Objects: geom_bar
```{r}
ggplot(data=exam_data, aes(x=RACE)) +
  geom_bar()
```

### 1.7.2 Geometric Objects: geom_dotplot
```{r}
ggplot(data=exam_data, aes(x=MATHS)) +
  geom_dotplot(binwidth=2.5, dotsize = 0.5) +
  scale_y_continuous(NULL, breaks = NULL)
```

### 1.7.3 Geometric Objects: geom_histogram()
```{r}
ggplot(data=exam_data, aes(x = MATHS)) +
  geom_histogram()       
```

### 1.7.4 Modifying a geometric object by changing geom()
```{r}
ggplot(data=exam_data, aes(x = MATHS)) +
  geom_histogram(bins=20, color="black", fill="light blue")
```

### 1.7.5 Modifying a geometric object by changing aes()
```{r}
ggplot(data=exam_data, aes(x = MATHS, fill = GENDER)) +
  geom_histogram(bins=20, color="grey30")
```
## (Alternate 1) Science and Gender Histogram

```{r}
ggplot(data=exam_data, aes(x = SCIENCE, fill = GENDER)) +
  geom_histogram(bins=15, color="grey30")
```

### 1.7.6 Geometric Objects: geom-density()
```{r}
ggplot(data=exam_data, aes(x = MATHS, colour = GENDER)) +
  geom_density()
```

### 1.7.7 Geometric Objects: geom_boxplot
```{r}
ggplot(data=exam_data, aes(y = MATHS, x = GENDER)) +
  geom_boxplot(notch=TRUE)
```

### 1.7.8 Geometric Objects: geom_violin

```{r}
ggplot(data=exam_data, aes(y = MATHS, x = GENDER)) +
  geom_violin()
```
## (Alternate 2) Science and Gender Violin plots

```{r}
ggplot(data=exam_data, aes(y = SCIENCE, x = GENDER)) +
  geom_violin()
```

### 1.7.9 Geometric Objects: geom_point()

```{r}
ggplot(data=exam_data, aes(x = MATHS, y = ENGLISH)) +
  geom_point()
```

### 1.7.10 geom objects can be combined

```{r}
ggplot(data=exam_data, aes(y = MATHS, x= GENDER)) +
  geom_boxplot() +                    
  geom_point(position="jitter", size = 0.5)        
```
## 1.8 Essential Grammatical Elements in ggplot2: stat
### 1.8.1 Working with stat()

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, x= GENDER)) +
  geom_boxplot()
```
### 1.8.2 Working with stat - the stat_summary() method

```{r}
ggplot(data=exam_data, aes(y = MATHS, x= GENDER)) +
  geom_boxplot() +
  stat_summary(geom = "point", fun = "mean",colour ="red", size=4)                        
```
## (Alternate 3) Math and Race Boxplots

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, x= RACE)) +
  geom_boxplot() +
  geom_point(stat = "summary", fun = "mean", colour ="red", size=3)
```
## (Alternate 4) Science and Race Boxplots

```{r}
ggplot(data=exam_data, 
       aes(y = SCIENCE, x= RACE)) +
  geom_boxplot() +
  geom_point(stat = "summary", fun = "mean", colour ="blue", size=3)               
```

### 1.8.3 Working with stat - the geom() method

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, x= GENDER)) +
  geom_boxplot() +
  geom_point(stat = "summary", fun = "mean", colour ="red", size=4)               
```
## (Alternate 5) English and Science Scatterplot with Best Fit curve

```{r}
ggplot(data=exam_data, 
       aes(x= ENGLISH, y=SCIENCE)) +
  geom_point() +
  geom_smooth(method=lm, linewidth=1)         
```

### 1.8.4 Adding a best fit curve on a scatterplot

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, linewidth=0.5)         
```
## 1.9 Essential Grammatical Elements in ggplot2: Facets
### 1.9.1 Working with facet_wrap()

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS)) +
  geom_histogram(bins=20) +
    facet_wrap(~ CLASS)
```

### 1.9.2 facet_grid() function

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS)) +
  geom_histogram(bins=20) +
    facet_grid(~ CLASS)
```

## 1.10 Essential Grammatical Elements in ggplot2: Coordinates
### 1.10.1 Working with Coordinate

```{r}
ggplot(data=exam_data, aes(x=RACE)) +
  geom_bar() +
  coord_flip()
```

### 1.10.2 Changing the y- and x-axis range

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, size=0.5) +
  coord_cartesian(xlim=c(0,100), ylim=c(0,100))
```
## 1.11 Essential Grammatical Elements in ggplot2: themes
### 1.11.1 Working with theme

```{r}
ggplot(data=exam_data, aes(x=RACE)) +
  geom_bar() +
  coord_flip() +
  theme_minimal()
```


## (Alternate 6) Density plot with Math
```{r}
ggplot(data=exam_data, aes(x = MATHS)) +
  geom_density(fill="#69b3a2", color="#e9ecef", alpha=0.8)
```
## (Alternate 7) Scatterplot of Science & Maths with geom_rug()
```{r}
ggplot(data=exam_data, aes(x = MATHS, y = SCIENCE)) +
  geom_point() +
  geom_rug(col="steelblue",alpha=0.1, size=1.5)
```