---
title: "Hands-on_Ex02"
author: "Andre Ong Jia Kang"
---

# 2.Beyond ggplot2 Fundamentals
```{r}
library(extrafont)
```
## Install and launching R packages

```{r}
pacman::p_load(ggrepel, patchwork, 
               ggthemes, hrbrthemes,
               tidyverse, ggiraph, plotly, DT) 
```

## Importing the data

```{r}
exam_data <- read_csv("data/Exam_data.csv")

```

## 2.3 Beyond ggplot2 Annotation: ggrepel 
###2.3.1 Working with ggrepel

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label_repel(aes(label = ID), 
                   fontface = "bold") +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```
### (Explore 1) Science vs English Scores with annotation

```{r}
ggplot(data=exam_data, 
       aes(x= SCIENCE, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label_repel(aes(label = ID), 
                   fontface = "bold") +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

## 2.4 Beyond ggplot2 Themes

```{r}
ggplot(data=exam_data, aes(x = MATHS)) +
  geom_histogram(bins=20, boundary = 100, color="grey25", fill="grey90") +
  theme_gray() +
  ggtitle("Distribution of Maths scores") 
```

### 2.4.1 Working with ggtheme package

```{r}
ggplot(data=exam_data, aes(x = MATHS)) +
  geom_histogram(bins=20, boundary = 100, color="grey25", fill="grey80") +
  ggtitle("Distribution of Maths scores") +
  theme_economist()
```
### (Explore 2) Science Histogram with theme

```{r}
ggplot(data=exam_data, aes(x = SCIENCE)) +
  geom_histogram(bins=20, boundary = 100, color="grey25", fill="grey80") +
  ggtitle("Distribution of Maths scores") +
  theme_economist()
```

### 2.4.2 Working with hrbthemes package

```{r}
ggplot(data=exam_data, aes(x = MATHS)) +
  geom_histogram(bins=20, boundary = 100, color="grey25", fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum()
```

## 2.5 Beyond Single Graph

```{r}
p1 <- ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") + 
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of Maths scores")

p2 <- ggplot(data=exam_data, 
             aes(x = ENGLISH)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of English scores")

p3 <- ggplot(data=exam_data, 
             aes(x= MATHS, 
                 y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

### 2.5.2 Combining two ggplot2 graphs

```{r}
p1 + p2
```
### (Explore 3) Combining 3 histogram graphs of english, math and science scores

```{r}
p4 <- ggplot(data=exam_data, 
             aes(x = SCIENCE)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of Science scores")

p1 + p2 + p4
```

### 2.5.3 Combining three ggplot2 graphs

```{r}
(p1 / p2) | p3
```
## (Alternate 1) Science and Math distribution, comparison with changes to size and cartesian coords

```{r}
p5 <- ggplot(data=exam_data, 
             aes(x= MATHS, 
                 y=SCIENCE)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=1.5) +  
  coord_cartesian(xlim=c(5,105),
                  ylim=c(5,105)) +
  ggtitle("Science scores versus Maths scores for Primary 3")

(p1 / p4) | p5
```

### 2.5.4 Creating a composite figure with tag

```{r}
((p1 / p2) | p3) + 
  plot_annotation(tag_levels = 'I')
```
## (Alternate 2) Creating a composite figure by using patchwork and ggtheme

```{r}
patchwork <- (p1 / p4) | p5
patchwork & theme_economist()
```
## (Alternate 3) Exploring tufte theme with patchwork
```{r}
patchwork <- (p1 / p2) | p3   
patchwork & theme_tufte()
```

### 2.5.5 Creating figure with insert

```{r}
p3 + inset_element(p2, 
                   left = 0.02, 
                   bottom = 0.7, 
                   right = 0.5, 
                   top = 1)
```

## (Alternate 4) Science vs Math Scatter figure with Science insert, changing theme and coords
```{r}
p5 + inset_element(p4, 
                   left = 0.01, 
                   bottom = 0.65, 
                   right = 0.5, 
                   top = 1)  + theme_stata()
```
